---
title: "Compare Business Dynamism Across the 50 States"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    source_code: embed
    vertical_layout: scroll
    orientation: rows
runtime: shiny
---

```{r setup, include=FALSE}
library(shiny)
library(flexdashboard)
library(ggplot2)
library(DT)
library(stringr)
library(dplyr)
library(tools)
library(plotly)

# Load file
busdyn <- read.csv(file = "BDS_Data_04_18.csv", fileEncoding="UTF-8-BOM")

# Remove Underscores in Column and Row Names
gsub(x = names(busdyn), pattern = "_", replacement = " ") 

```
Inputs {.sidebar}
-----------------------------------------------------------------------
```{r sidebar}
# User selects which state(s) to graph------------------------------------------
checkboxGroupInput("state", "State(s)", choices = unique(busdyn$State), inline = TRUE
                  )

# User selects which year to graph-------------------------------------------
sliderInput("year", "Year(s)",
            min = (2004), max = (2018),
            value = c(2018), sep = ""
            )

# User selects type of plot for third visualization---------------------------------------------
radioButtons("plot_type", "Choose plot type showing job reallocation across all states", 
             choices = c("Box plot", "Violin Plot"), selected = c("Box plot"))
```
Outputs


Row {data-height=175}
-----------------------------------------------------------------------
### Introduction

This dashboard measures American business dynamism--a crucial input for job creation. All data is sourced from the U.S. Census Business Dynamics Statistics Survey.

Chart 1 measures business creation rate, or the share of all firms that are 1 year old or younger. Chart 2 measures business creation rate against job creation rate from these new businesses, looking at where young companies are able to quickly grow in their first year. Chart 3 measures job reallocation rate, the sum of all jobs created and destroyed - looking at how quickly jobs move from shrinking to expanding firms.


Row {data-height=150}
-----------------------------------------------------------------------
### Average Annual Rate of Business Creation (selected data)

```{r valueboxmeanbiz}
# Value Box showing average value of Business Creation Rate
renderValueBox({
valueBox(color = "gold", value = round(mean(busdyn_subset()$Rate_of_establishments_born_prev_year), 1), "Average Annual Rate of Business Creation (selected data)", icon = NULL)   })

```

### Max. Annual Rate of Business Creation (selected data)

```{r valueboxmeanjobs}
# Value Box showing average value of Job Creation Rate
renderValueBox({
valueBox(color = "gold", value = round(mean(busdyn_subset()$Rate_of_jobs_created_from_opening_establishments_prev_year), 1), "Average Annual Rate of Job Creation at Young Firms (selected data)", icon = NULL) 
  })

```

### Min. Annual Rate of Business Creation (selected data)

```{r valueboxmeanreallocation}
# Value Box showing average value of Job Reallocation Rate
renderValueBox({
valueBox(color = "gold", value = round(mean(busdyn_subset()$Rate_of_reallocation_prev_year, 1)), "Mean Annual Rate of Job Reallocation (selected data)", icon = NULL) 
  })
```

Row {data-height=650, .tabset}
-----------------------------------------------------------------------
### Chart 1: Rate of Business Creation for Selected States & Year

```{r first plot}
# Create subset of dataset based on user's year and state inputs 
busdyn_subset <- reactive({
   if (length(input$state) > 0 ) { #select all states if user did not select any
     subset (busdyn, State == input$state & Year == input$year, select = c(State, Year, Rate_of_establishments_born_prev_year, Rate_of_jobs_created_from_opening_establishments_prev_year, Rate_of_reallocation_prev_year))
   }
    else{subset (busdyn, Year == input$year, select = c(State, Year, Rate_of_establishments_born_prev_year, Rate_of_jobs_created_from_opening_establishments_prev_year, Rate_of_reallocation_prev_year))}
  })


# Create histogram of data subset for first visualization
renderPlotly({
ggplotly(ggplot(busdyn_subset(), aes(x = State, y = Rate_of_establishments_born_prev_year)) +
           theme(axis.text.x=element_text(angle=60,hjust=1)) +
         geom_bar(stat = "identity", width = .4, fill = "blue4") +
  labs(y = "Rate of Establishments Born in the Previous Year",
       x = "State",
       title ="Rate of Business and Job Creation for Selected States & Year",
       subtitle = "Share of all firms that are 1 year old or younger"),
         tooltip = c("x", "y"))
})
```

### Chart 2: Rate of Business and Job Creation for Selected States & Year

```{r second plot}

# Create scatterplot of data subset for second visualization
renderPlotly({
ggplotly(ggplot(busdyn_subset(), aes(x = Rate_of_establishments_born_prev_year,
                            y = Rate_of_jobs_created_from_opening_establishments_prev_year, text = State)) + 
         geom_point(aes(col = State)) +
         labs(y = "Rate of Jobs Created from New Establishments",
              x = "Rate of Establishments Born Previous Year",
              title = "Rate of Business and Job Creation for Selected States & Year",
              subtitle = "Share of all firms that are 1 year old or younger and share of jobs at these firms"),
         tooltip = c("text", "x", "y") 
)
})
```

### Chart 3: Overall Rate of Business Dynamism for all 50 States

```{r third plot}
# Create boxplot or violin plot of all data based on user's choice of plot
renderPlotly({
  if(input$plot_type == "Box Plot"){
    (ggplotly(ggplot(busdyn, aes(x = Year, y = Rate_of_reallocation_prev_year)) +
               geom_violin()+
               labs(y = "Annual Rate of Job Reallocation",
                    x =  "Year",
                    title = "The Distribution of Job Reallocation Rates across the States, 2004-2018",
                    subtitle = "Sum of jobs created and destroyed - measuring how quickly jobs more from shrinking to expanding firms",
                    caption = "Source: U.S. Census Business Dynamics Statistics Survey")
          ))}
    
  else {
    (ggplotly(ggplot(busdyn, aes(x = Year, y = Rate_of_reallocation_prev_year, text = State)) +
      geom_boxplot(fill = "blue4")+
        scale_x_discrete(breaks=NULL) +
      labs(y = "Annual Rate of Job Reallocation",
           x =  "Year",
           title = "The Distribution of Job Reallocation Rates across the States, 2004-2018",
           subtitle = "Sum of jobs created and destroyed - measuring how quickly jobs more from shrinking to expanding firms"),
      tooltip = "text"))
}
})
```

### Data Table

```{r datatable}
# Create data table corresponding to selected data
renderDataTable(busdyn_subset(), options = list(pageLength = 10, autoWidth = TRUE, columnDefs = list(list(width = "8px", targets = "all"()))), 
                    rownames = FALSE)
```